Matching Bullets

### 2015-11-23 imaging-paper.Rnw
bit of editing

It's clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.

@@

\begin{figure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residuala}Residual structure of a surface crosscut at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residualb} Residual structure of a surface crosscut at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of  circular fits at two different cross-section. The residual structure in (b) shows systematic structure, indicating that the fit is not entirely appropriate. }
\end{figure}

A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in vertical direction. We therefore fit a circle for each cross section of the bullet. Figure \ref{fig:circlefits} illustrates an attempt at matching one of the six lands to the first part of bullet 1, which is shown as a black line. There is some global structure evident that makes focusing on the deviations a bit challenging.

@@

<<circlefits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.', fig.pos='H'>>=
qplot(y, resid, data=subset(db2, x <= 80*1.5625^2 & x >=75*1.5625^2), colour=factor(x), geom="line", size=I(1)) + 
  facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group = x), colour="black", size=1, alpha=0.2, 
             data = filter(db1, part==1, x <= 80*1.5625^2, x >=75*1.5625^2)[,c("y", "resid", "x")]) + ggtitle("Part 1 of Bullet 1 in black")
@

@@

<<circlefits2, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 2 of bullet 1 overlayed.', fig.pos='H', cache=TRUE>>=
qplot(y, resid, data=subset(db2, x <= 80*1.5625^2 & x >=75*1.5625^2), colour=factor(x), geom="line", size=I(1)) + 
  facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group =x), colour="black", size=1, alpha=0.2, 
             data = filter(db1, part==2, x <= 80*1.5625^2, x >=75*1.5625^2)[,c("y", "resid", "x")]) + ggtitle("Part 2 of Bullet 1 in black")
@

@@

Next, we fit a Loess regression to the data

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}
\centering
\begin{subfigure}[t]{\linewidth}
\caption{Crosscuts at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line", alpha=I(0.6)) + 
  scale_colour_brewer("", palette="Set1") + theme_bw() +
  theme(legend.position = c(.9, .2))
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
      geom="line", alpha=I(0.6)) + 
  scale_colour_brewer(palette="Set1") + theme_bw() +
  theme(legend.position="none")
@
    }
\end{subfigure}
\caption{\label{fig:cross100} Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}
\centering
    \begin{subfigure}[t]{\linewidth}
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.49\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=bullet, size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey30", size=0.5) +
  facet_grid(bullet~.) +
  scale_colour_brewer("", palette="Set1") +
  theme_bw() +
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Values above/below a threshold level are defined as peaks and grooves. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='.49\\textwidth', warning=FALSE>>=
qplot(data=lofX, x=y, y=l30, colour=type, geom="line", 
      group=bullet, size = I(0.75)) + 
  facet_grid(bullet~.) +
  ylab("Individual characteristics") + 
#  geom_point(aes(y=r05)) +
  geom_rect(xmin = 200, xmax = 2260, ymin = -.75, ymax = 0.75, colour=NA, fill=alpha("grey90", alpha =0.5)) +
  scale_colour_brewer(palette="Set2", na.value = NA) +
  theme_bw() +
  theme(legend.position="bottom")
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Grey backgrounds identify a line on one of the bullets. Overlaps of the same color indicate matching striation marks.
    Matching striations are marked by an `o', mismatches are marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.49\\textwidth'>>=
ggplot() +
  geom_rect(aes(xmin = miny, xmax = maxy), ymin = 0.25, ymax=2.75, fill="grey80", data = subset(lines, lineid != 0)) +
  geom_tile(data=lofX, aes(x=y, y=bullet, fill=type)) + 
  scale_y_discrete(expand = c(0.2,0)) +
  scale_fill_brewer(palette="Set2") + 
  theme_bw() +
  ylab("") +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= .4, label= "x", data = subset(lines, !match & lineid !=0)) +
  geom_text(aes(x = meany), y= .4, label= "o", data = subset(lines, match & lineid !=0))
@
    }
\end{subfigure}
\caption{\label{fig:match} Matching striation marks: smooth (a), cut (b), and match (c).}
\end{figure}

>>>>

It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.

@@

\begin{figure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residuala}Residual structure of a surface crosscut at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residualb} Residual structure of a surface crosscut at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of  circular fits at two different cross-section. Both residual plots show systematic structures, indicating that the fit is not entirely appropriate. }
\end{figure}

A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in vertical direction. We therefore fit a circle for each cross section of the bullet. Figure~\ref{fig:circlefits} illustrates an attempt at matching one of the six lands to the first part of bullet 1, which is shown as a black line. There is some global structure evident that makes focusing on the deviations a bit challenging.

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[b]{\linewidth}
\caption{Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.\label{fig:circlefits}}{%
<<circlefits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
      geom="line", size=I(1), colour=I("grey30")) + 
  facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group = x), colour="black", size=1, alpha=0.2, 
             data = filter(db1, part==1, x == 100)[,c("y", "resid", "x")]) + ggtitle("Part 1 of Bullet 1 in black")
@
}
\end{subfigure}    
\begin{subfigure}[b]{\linewidth}
    \caption{Circle fit for each cross section of bullet 2, with part 2 of bullet 1 overlayed.\label{fig:circlefits2}}{%
<<circlefits2, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE,  cache=TRUE>>=
qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
      geom="line", size=I(1), colour=I("grey30")) + 
  facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group =x), colour="black", size=1, alpha=0.2, 
             data = filter(db1, part==2, x == 100)[,c("y", "resid", "x")]) + ggtitle("Part 2 of Bullet 1 in black")
@    
}
\end{subfigure}
\caption{An example of a scan of a bullet land from one groove to the next. }
\end{figure}

@@

@@

Next, we fit a Loess regression to the data.

\hh{XXX could you include a series of plots here, walking through the individual steps of this list? I like the ones that you've already got in the loess fit function ... }

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}
\centering
\begin{subfigure}[t]{\linewidth}
\caption{Crosscuts at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line", alpha=I(0.6)) + 
  scale_colour_brewer("", palette="Set1") + theme_bw() +
  theme(legend.position = c(.9, .2))
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
      geom="line", alpha=I(0.6)) + 
  scale_colour_brewer(palette="Set1") + theme_bw() +
  theme(legend.position="none")
@
    }
\end{subfigure}
\caption{\label{fig:cross100} Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}
\centering
    \begin{subfigure}[t]{\linewidth}
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=bullet, size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey30", size=0.5) +
  facet_grid(bullet~.) +
  scale_colour_brewer("", palette="Set1") +
  theme_bw() +
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Values above/below a threshold level are defined as peaks and grooves. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='\\textwidth', warning=FALSE>>=
qplot(data=lofX, x=y, y=l30, colour=type, geom="line", 
      group=bullet, size = I(0.75)) + 
  facet_grid(bullet~.) +
  ylab("Individual characteristics") + 
#  geom_point(aes(y=r05)) +
  geom_rect(xmin = 200, xmax = 2260, ymin = -.75, ymax = 0.75, colour=NA, fill=alpha("grey90", alpha =0.5)) +
  scale_colour_brewer(palette="Set2", na.value = NA) +
  theme_bw() +
  theme(legend.position="bottom")
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Grey backgrounds identify a line on one of the bullets. Overlaps of the same color indicate matching striation marks.
    Matching striations are marked by an `o', mismatches are marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='\\textwidth'>>=
ggplot() +
  geom_rect(aes(xmin = miny, xmax = maxy), ymin = 0.25, ymax=2.75, fill="grey80", data = subset(lines, lineid != 0)) +
  geom_tile(data=lofX, aes(x=y, y=bullet, fill=type)) + 
  scale_y_discrete(expand = c(0.2,0)) +
  scale_fill_brewer(palette="Set2") + 
  theme_bw() +
  ylab("") +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= .4, label= "x", data = subset(lines, !match & lineid !=0)) +
  geom_text(aes(x = meany), y= .4, label= "o", data = subset(lines, match & lineid !=0))
@
    }
\end{subfigure}
\caption{\label{fig:match} Matching striation marks: smooth (a), cut (b), and match (c).}
\end{figure}
